---
title: "Data Analysis"
output: 
  html_document:
    toc: true
    toc_float: true
---

```{r, include=FALSE}
library(tidyverse)
library(modelr)

set.seed(1)
museum_edu_df <- read.csv("museum_edu_data.csv")
```

# Is the number of people who hold college degrees significantly related to the amount of museums in their zip code?

```{r}
# Count the number of museums per zip code
museum_count_per_zip <- museum_edu_df |>
  group_by(zip_code) |>
  summarize(n_museums = n())

# Merge the museum count back into the main dataset
museum_edu_df <- left_join(museum_edu_df, museum_count_per_zip, by = c("zip_code"), suffix = c(".museum_count_per_zip", ".museum_edu_df"))

# Linear regression: Number of Museums vs Highest Educational Attainment Category
regression_model_bach <- lm(bachelors ~ n_museums, data = museum_edu_df)

# Creating Regression Model Table
regression_model_bach |>
  broom::tidy() |>
  select(term, estimate, p.value) |>
  knitr::kable(digits=10)
```

E(bachelors) = β0 + β1(museum_n)

Intercept (Estimated Constant):
The intercept is 10.39586. This is the estimated percentage of individuals with a bachelor's degree when the number of museums (n_museums) is zero.

Coefficient for n_museums (Number of Museums):
The coefficient for n_museums is 0.17138. This represents the estimated change in the percentage of individuals with a bachelor's degree for each additional museum. The positive sign indicates a positive relationship, suggesting that, on average, as the number of museums increases, the percentage of individuals with a bachelor's degree also tends to increase.

Statistical Significance:
The associated p-value is very small (1.81e-07). This indicates that the number of museums is statistically significant in predicting the percentage of individuals with a bachelor's degree.

In summary, the results suggest a statistically significant, positive relationship between the number of museums and the percentage of individuals with a bachelor's degree.


# Is the median income significantly related to the amount of museums in a zip code?

```{r}
#Linear regression model: museums and medium income
regression_model_income <- lm(n_museums ~ median_income, data = museum_edu_df)

# Creating Regression Model Table
regression_model_income |>
  broom::tidy() |>
  select(term, estimate, p.value) |>
  knitr::kable(digits=10)
```

E(n_museums) = β0 + β1(median_income)

Intercept (Estimated Constant):

The intercept is 2.538. This is the predicted number of museums when the median income is zero. 

Coefficient for median_income:
The coefficient for median_income is -0.000014379. This represents the estimated change in number of museums for each unit change in income. The negative sign indicates a negative relationship, suggesting that, on average, as the number of median income increases, the number of museums tends to decrease.

Statistical Significance:
The associated p-value is very small (2.73e-05). This indicates that the number of museums is statistically significant in predicting median income.

In summary, the results suggest a statistically significant,negative relationship between the median income and number of museums. 

#Is median income significantly related to the number of people who hold bachelor's degrees?
```{r}
#Linear regression model: college degrees and medium income
regression_model_bachincome <- lm(bachelors ~ median_income, data = museum_edu_df)

# Creating Regression Model Table
regression_model_bachincome |>
  broom::tidy() |>
  select(term, estimate, p.value) |>
  knitr::kable(digits=10)
```

E(bachelors) = β0 + β1(median_income)

Intercept (Estimated Constant):

The intercept is -6.725. This is the estimated percentage of individuals with a bachelor's degree when the median income is zero.

Coefficient for median_income:
The coefficient for median_income is 0.0004709779 This represents the estimated change in estimated percentage of individuals with a bachelor's degree for each unit change in income. The positive sign indicates a positive relationship, suggesting that, on average, as the number of museums increases, the median income tends to increase

Statistical Significance:
The associated p-value is very small (2.73e-05). This indicates that the median income is statistically significant in percentage of individual's with a bachelor's degrees.

In summary, the results suggest a statistically significant, positive relationship between the number of bachelor's degrees and median income.

# Is median income a confounder on the relationship between educational attainment and the amount of museums?
since median income is associated with both the number of museums and percentage of individuals with a bachelor's degree, we should investigate if median income is a confounder between bachelor's degrees and number of museums. 

```{r}
#linear regression model testing for confounding
model4 = lm(formula = bachelors ~ median_income + n_museums, data = museum_edu_df)

# Creating Regression Model Table
model4 |>
  broom::tidy() |>
  select(term, estimate, p.value) |>
  knitr::kable(digits=10)
```

E(bachelors) = β0 + β1(median_income) + β2(n_museums)

Intercept (Estimated Constant):
The intercept is -7.350. This represents the estimated percentage of individuals with a bachelor's degree when the number of museums (n_museums) is zero and median income (median_income) is zero.

Statistical significance: 
The p-values associated with each coefficient are very small (<2×10^−16), indicating that both median_income and n_museums are statistically significant predictors of the response variable.

Coefficient for n_museums:
The coefficient for n_museums is 0.245. his represents the estimated change in the percentage of individuals with a bachelor's degree for each additional museum, adjusted for median income. 
In the unadjusted model, the coefficient was: 0.17138. There is a 30% difference between the crude and unadjusted models, which is greater than 10%, signifying that income does infact confound the relationship between high educational attainment and amount of museums in a zipcode.

## Comparison of models including income and excluding income
```{r}
cv1_df = 
  crossv_mc(museum_edu_df, 100) 

cv1_df |> 
  mutate(
    mod_1  = map(train, \(df) lm(bachelors ~ n_museums, data = museum_edu_df)),
    mod_2  = map(train, \(df) lm(bachelors ~ median_income + n_museums, data = museum_edu_df))) |> 
  mutate(
    rmse_1 = map2_dbl(mod_1, test, \(mod, df) rmse(model = mod, data = df)),
    rmse_2 = map2_dbl(mod_2, test, \(mod, df) rmse(model = mod, data = df))) |>
  select(starts_with("rmse")) |>
  pivot_longer(
    everything(), names_to="model", values_to="rmse", names_prefix="rmse_") |>
  ggplot(aes(x=model, y=rmse)) + geom_violin()
```

The model excluding income has a much higher root mean squared errors (RMSE) value than the model including income. This means that the model including income better fits our data.

# Testing for Interaction between Median Income and Number of Museums
```{r}
#linear regression model testing for confounding
model_int = lm(formula = bachelors ~ median_income + n_museums + median_income*n_museums, data = museum_edu_df)

# Creating Regression Model Table
model_int |>
  broom::tidy() |>
  select(term, estimate, p.value) |>
  knitr::kable(digits=10)
```

E(bachelors) = β0 + β1(median_income) + β2(n_museums) + β3(median_income*n_museums)

The significant interaction term in the model implies that the effect of n_museums on bachelors is not constant across all levels of median_income. In other words, the relationship between the number of museums (n_museums) and the likelihood of having a bachelor's degree (bachelors) is influenced by the level of median_income.

## Comparison of income including models with and without interaction term:
```{r}
cv2_df = 
  crossv_mc(museum_edu_df, 100) 

cv2_df |> 
  mutate(
    mod_1  = map(train, \(df) lm(formula = bachelors ~ median_income + n_museums + median_income*n_museums, data = museum_edu_df)),
    mod_2  = map(train, \(df) lm(formula = bachelors ~ median_income + n_museums, data = museum_edu_df))) |> 
  mutate(
    rmse_1 = map2_dbl(mod_1, test, \(mod, df) rmse(model = mod, data = df)),
    rmse_2 = map2_dbl(mod_2, test, \(mod, df) rmse(model = mod, data = df))) |>
  select(starts_with("rmse")) |>
  pivot_longer(
    everything(), names_to="model", values_to="rmse", names_prefix="rmse_") |>
  ggplot(aes(x=model, y=rmse)) + geom_violin()
```

The model with an interaction term between bachelors and median income has a slightly lower RMSE value than the model without an interaction term. Ultimately, the model with interaction terms appears to be the best model.
