---
title: "Exploratory Data Analysis"
---

```{r, include=FALSE}
library(tidyverse)
library(plotly)
```

```{r, include=FALSE}
#data import and cleaning

museum_df = 
  read_csv("./Data/museums.csv", na = c("NA")) %>% 
  janitor::clean_names() %>% 
  select(
    -legal_name, -alternate_name, -institution_name, -phone_number, -employer_id_number,
    -ends_with("administrative_location"), -starts_with("street_address")) |> 
  drop_na(zip_code_physical_location) |> 
  mutate(
    city = city_physical_location,
    state = state_physical_location,
    zip_code = zip_code_physical_location
  ) |> 
  select(-ends_with("physical_location"))

education_df = 
  read_csv(
    "./Data/ACSST5Y2014.S1501-Data.csv", skip = 1, na = c("-", "**", "***", "(x)")) %>% 
  janitor::clean_names() |> 
  mutate(
    zip_code = as.numeric(str_replace(geographic_area_name, "ZCTA5 ", "")),
    less_than_HS = total_estimate_less_than_high_school_graduate,
    HS_diploma = total_estimate_high_school_graduate_includes_equivalency,
    some_college = total_estimate_some_college_or_associates_degree,
    bachelors = total_estimate_bachelors_degree_or_higher,
    median_income = total_estimate_median_earnings_in_the_past_12_months_in_2014_inflation_adjusted_dollars_population_25_years_and_over_with_earnings
  ) |> 
  select(zip_code, less_than_HS, HS_diploma, some_college, bachelors, median_income)

museum_edu_df =
  inner_join(museum_df, education_df)
```

### U.S. States by Number of Museums/Zoos/Aquariums

```{r, echo = FALSE}
museum_edu_df |> 
  count(state) |> 
  mutate(state = fct_reorder(as_factor(state), n)) |> 
  plot_ly(x = ~state, y = ~n, color = ~state, type = "bar", colors = "viridis")
```

New York state has far and away the most museums, zoos, and aquariums (n = 609). DC also has a lot when you consider the fact that it is essentially just a city (n = 42), but this makes sense when you think about DC as the capitol city of the whole country.

### Number of Museums in a City by Average Median Income of the Population

```{r, echo = FALSE, warning = FALSE}
museum_edu_df |> 
  group_by(city) |> 
  summarize(
    n_museums = n(),
    avg_median_income = mean(median_income)
  ) |> 
  filter(n_museums >= 5) |> 
  plot_ly(x = ~n_museums, y = ~avg_median_income, type = "scatter", mode = "markers",
          text = ~city)
```

The above scatterplot shows the number of museums in a city by the average median income of the population over 25 years. Only cities that have at least five museums are included. There does not appear to be much of a correlation, between the two variables; we can see this with the large amount of variation in income for the cities with 5 museums. Notably, New York City has 68 museums, and an average median annual income of $68.7K. Average median income is obtained by average the median income by zip codes within the city.


